---
title: "Eval_Lab"
author: "Brian Wright"
date: "10/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Throughout your early career as a data scientist you've built article summaries, 
explored NBA talent, minded text on climate change news but you've suddenly realized you need to enhance your ability to assess the models you are building. 
As the most important part about understanding any machine learning model 
(or any model, really) is understanding it's weakness and or vulnerabilities. 

In doing so you've decided to practice on a dataset that are of interest to you, and use a approach to which you are familiar, kNN. 

Part 1. Select either as a lab group or as individuals a dataset that is of interest to you/group. Define a question that can be answered using classification, specifically kNN, for the dataset. 

### Can we use patient health data to predict whether or not patients are at risk of death?

Part 2. In consideration of all the metrics we discussed what are a few key metrics that should be tracked given the question you are working to solve?

### F1-score
  In our problem, F1-score would be a useful metric to track given the concern that False Positive and False Negatives pose. The cost of a false positive and false negative are both very high. A false positive would mean predicting a patient would die when they actually would not. A false negative would mean predicting a patient would not die when they actually would. The first case would lead to a lot of unnecessary distress while the second case may give false security. 
### Sensitivity/Recall
  Since the cost of false negatives is highest, it's best to focus on track recall. We want to be able to predict as many actual positives as possible.
### Prevalence
  We should focus on prevalance so as to be able to accurately evaluate the performance of our model. We want to avoid the case where we think our model performs well, but our data only contains a small percentage of the positive class. 

Part 3. Build a kNN model and evaluate the model using the metrics discussed in class (Accuracy, TPR, FPR, F1, Kappa, LogLoss and ROC/AUC). Make sure to calculate the baserate or prevalence to provide a reference for some of these measures. Even though you are generating many of the metrics we discussed, summarize the output of the key metrics you established in part 2. 

Part 4.  Consider where miss-classification errors are occurring, is there a pattern? If so discuss this pattern and why you think this is the case. 

  When miss-classfication occurs, our model tends to make the probability of the incorrect class very high. This is why the log-loss of our model is so high. This might be due to the use of binary variables when training our model.

Part 5. Based on your exploration in Part 3, change the threshold using the function provided in the in-class example, what differences do you see in the evaluation metrics? Speak specifically to the metrics that are best suited to address the question you are trying to answer. 

  When making the threshold higher, the precision of the model increases. However, this reduces the recall/sensitivity of the model. Likewise, when the threshold is made lower, the model tends to predict the positive class more frequently. This increases recall/sensitivity but decreases precision. 

Part 6. Summarize your findings focusing on speaking through your question, what does the evaluation outputs mean when answering the question you've proposed?

  According to our experiments, we have determined that we can somewhat accurately predict whether or not a patient will die based off various attributes of that patient. The accuracy of the best k we had was 0.75. More work must be done to improve the recall of our model. The costs of false negatives is very high, since it can result in an unexpected death of a patient. To improve our model, we would recommend gathering more data, adjusting our thresholds, and conducting more research in the problem domain.

Recommendations for improvement might include gathering more data, adjusting the threshold, adding new features, changing your questions or maybe that it's working fine at the current level and nothing should be done.

Regardless of the outcome, what should we be aware of when your model is deployed (online versus offline)? 
  
  We should be aware of the fact that our model relies on a fixed set of data. New-incoming data can make the model useless or obsolete. we want to be able to actively monitor it so that we can make adjustments if necessary. 
  

---
title: "Eval_Lab"
author: "Brian Wright"
date: "10/25/2020"
output: html_document
---

```{r}
library(tidyverse)
library(psych)
library(mltools)
library(data.table)
library(caret)
library(class)
```

```{r}
heart_data <- read_csv("heart_failure_clinical_records_dataset.csv")
heart_data$age <- scale(heart_data["age"], center = TRUE, scale = TRUE)
heart_data$creatinine_phosphokinase <- scale(heart_data["creatinine_phosphokinase"], center = TRUE, scale = TRUE)
heart_data$ejection_fraction <- scale(heart_data["ejection_fraction"], center = TRUE, scale = TRUE)
heart_data$platelets <- scale(heart_data["platelets"], center = TRUE, scale = TRUE)
heart_data$serum_creatinine <- scale(heart_data["serum_creatinine"], center = TRUE, scale = TRUE)
heart_data$serum_sodium <- scale(heart_data["serum_sodium"], center = TRUE, scale = TRUE)
heart_data$time <- scale(heart_data["time"], center = TRUE, scale = TRUE)
#View(heart_data)
```

```{r}
# Let's run the kNN algorithm on our heart data. 
# Check the composition of labels in the data set. 
table(heart_data$`DEATH_EVENT`)[2] / sum(table(heart_data$`DEATH_EVENT`))
# This means that at random, we have an 32.1% chance of correctly picking
# out a dead individual. Let's see if kNN can do any better.
part_index_1 <- createDataPartition(heart_data$`DEATH_EVENT`,
                                           times=1,
                                           p = 0.70,
                                           groups=1,
                                           list=FALSE)
#View(part_index_1)
train_heart <- heart_data[part_index_1,]
tune_and_test <- heart_data[-part_index_1, ]
#The we need to use the function again to create the tuning set 
tune_and_test_index <- createDataPartition(tune_and_test$`DEATH_EVENT`,
                                           p = .5,
                                           list = FALSE,
                                           times = 1)
tune_heart <- tune_and_test[tune_and_test_index, ]
test_heart <- tune_and_test[-tune_and_test_index, ]
dim(train_heart)
dim(tune_heart)
dim(test_heart)
```
## Train the classifier 

```{r}
# Let's train the classifier for k = 3 using the class package. 
# k-Nearest Neighbor is a randomized algorithm, so make sure to
# use set.seed() to make your results repeatable.
set.seed(1982)

trctrl <- trainControl(method = "repeatedcv",
                       number = 10,
                       repeats = 3) # generic control to pass back into the knn mode using the cross validation method. 

death_knn <- train(as.factor(DEATH_EVENT)~.,
                  data = train_heart[, c("age", "creatinine_phosphokinase", "ejection_fraction", "platelets", "serum_creatinine", "serum_sodium", "time", "high_blood_pressure", "diabetes", "sex", "smoking", "anaemia", "DEATH_EVENT")],
                  method="knn",
                  tuneLength=10,
                  trControl= trctrl,#cv method above, will select the optimal K
                  preProcess="scale") #already did this but helpful reference

death_knn
```
Now let's evaluate our model and replicate some of the evaluation metrics we've been discussing. 
```{r}
#First we need to do some predictions using the test data 

death_eval <-(predict(death_knn,newdata = test_heart))#generates 1s and 0s

View(death_eval)

#predict is polymorphic, meaning it doesn't care about the type of model passed into it
death_eval_prob <- predict(death_knn,newdata = test_heart, type = "prob")#this gives us the predicted prob, we will need these later for the fairness evaluation
View(death_eval_prob)

View(test_heart$DEATH_EVENT)
table(death_eval, test_heart$DEATH_EVENT)#essential the confusion matrix, though we can make a fancy one using caret built in functions


confusionMatrix(death_eval, as.factor(test_heart$DEATH_EVENT), positive = "1", dnn=c("Prediction", "Actual"), mode = "sens_spec")#change to everything

#Quick function to explore various threshold levels and output a confusion matrix

adjust_thres <- function(x, y, z) {
  #x=pred_probablities, y=threshold, z=test_DEATH_EVENT
  thres <- as.factor(ifelse(x > y, 1,0))
  confusionMatrix(thres, z, positive = "1", dnn=c("Prediction", "Actual"), mode = "everything")
}


adjust_thres(death_eval_prob$`1`,.60, as.factor(test_heart$DEATH_EVENT)) #Not much changes here because of the high probability splits of the data DEATH_EVENTs. Let's take a closer look. We can see that the algo isn't marginally mis-classifying these rows, it's very confidently wrong. Which likely means that there's too much emphasis being placed on too small a number of variables, principally the funfetti variable. 

death_eval_prob$test_heart <- as.factor(test_heart$DEATH_EVENT)

#notice that this model is probably relying too much on a single variable (fun-fetti) b/c distribution of probabilities is so consistent (underfitting)
View(death_eval_prob)

(error = mean(death_eval != as.factor(test_heart$DEATH_EVENT)))#overall error rate, on average when does our prediction not match the actual, looks like around 15%, really just ok. 

```

ROC/AUC
```{r}
#In order to use most evaluation packages it's just easier to have are predictions and targets in one place. 
death_eval <- tibble(pred_class=death_eval, pred_prob=death_eval_prob$`1`,target=as.numeric(test_heart$DEATH_EVENT))
str(death_eval)
pred <- prediction(death_eval$pred_prob,death_eval$target)
View(pred)
tree_perf <- performance(pred,"tpr","fpr")
plot(tree_perf, colorize=TRUE)
abline(a=0, b= 1)
tree_perf_AUC <- performance(pred,"auc")
print(tree_perf_AUC@y.values)
```

LogLoss
```{r}
View(death_eval)
LogLoss(as.numeric(death_eval$pred_prob), as.numeric(test_heart$DEATH_EVENT))
#We want this number to be rather close to 0, so this is a pretty terrible result. 
```


F1 Score 
```{r}
pred_1 <- ifelse(death_eval_prob$`1` < 0.5, 0, 1)
View(pred_1)
F1_Score(y_pred = pred_1, y_true = death_eval_prob$test_heart, positive = "1")
```



